---
title: "Loan Payback Prediction"
author: "May Tran"
date: "May 29, 2018"
output: word_document
---

```{r}
library(car)
library(tidyverse)
library(caret)
library(GGally)
library(ROCR)
library(ggplot2)
library(RcmdrMisc)
library(caTools)
loans <- read.csv(".../loan_data.csv")
summary(loans)
```

# Investigate data types of the columns

```{r}
str(loans)
```

# Transfer the data type to correct format

```{r}
loans$days.with.cr.line <- as.numeric(loans$days.with.cr.line)
loans$delinq.2yrs <- as.numeric(loans$delinq.2yrs)
loans$inq.last.6mths <- as.numeric(loans$inq.last.6mths)
loans$long_term <- as.numeric(loans$long_term)
```

# Understand data: use ggplot to visualize

```{r}
ggplot(data = loans,aes(fico)) + geom_histogram(aes(fill=as.factor(loans$notfullypaid)), color = 'black') + theme_bw()
```

The higher fico is, the less likely the loan is default 


```{r}
ggplot(data = loans, aes(fico)) + geom_histogram(aes(fill = as.factor(credit.policy)), color = 'black') + theme_bw()
```

The higher the fico is, the more likely the loan is accepted by the credit policy. 


```{r}
ggplot(data = loans,aes(dti)) + geom_histogram(aes(fill = as.factor(notfullypaid)),color = "black") +theme_bw()
```

```{r}
ggplot(data = loans, aes(int.rate,fico)) + geom_point() + stat_smooth()
ggplot(data = loans, aes(int.rate)) + geom_histogram(aes(fill = purpose), color = "black")
ggplot(data = loans, aes(int.rate)) + geom_histogram(aes(fill = as.factor(notfullypaid)), color = "black")

```
The higher fico is the lower the interest is
Interest rate seems irrelevant to loan purpose and loan payment ability

```{r}
set.seed(123)
sample <- sample.int(n = nrow(loans), size = floor(.9*nrow(loans)), replace = F)
loanstrain <- loans[sample,]
loanstest <- loans[-sample,]
```

# Model 1: How purpose relates to loan payment

```{r}
mod1 <- glm(notfullypaid ~ purpose, data = loanstrain)
summary(mod1)
mod1rsquared <- (mod1$null.deviance- mod1$deviance)/mod1$null.deviance
mod1rsquared
```

# Model 2: How installment amount, annual income, and interest rate relate to loan payment ability

```{r}
mod2 <- glm(notfullypaid ~ installment + log.annual.inc + int.rate, data = loanstrain)
summary(mod2)
mod2rsquared <- (mod2$null.deviance- mod2$deviance)/mod2$null.deviance
mod2rsquared
```


```{r}
str(loanstrain)
```

# Model 3: How all variables impact the loan payment ability

```{r}
mod3 <- glm(notfullypaid ~ ., data = loanstrain)
summary(mod3)
mod3rsquared <- (mod3$null.deviance- mod3$deviance)/mod3$null.deviance
mod3rsquared
```

# Reduced Model 3: Use stepwise to remove insignificant variables from Model 3

```{r}
redmod3 <- stepwise(mod3)
summary(redmod3)
redmod3rsquared <- (redmod3$null.deviance- redmod3$deviance)/redmod3$null.deviance
redmod3rsquared
```

# Model 4: Manually reduced model: From Reduced model 3, select statistically significant variables

```{r}
mod4 <- glm(notfullypaid ~ credit.policy + installment + log.annual.inc + fico + revol.bal + inq.last.6mths, data = loanstrain)
summary(mod4)
mod4rsquared <- (mod4$null.deviance- mod4$deviance)/mod4$null.deviance
mod4rsquared

```

# Final Model: Reduced Model 3 is chosen to predict the probability of not fully payment

```{r}
loanstrain$Mod3PredTrain <- predict(redmod3, newdata = loanstrain, type = "response")
loanstrain$Mod3ClassTrain <- ifelse(loanstrain$Mod3PredTrain > 0.5, 1,0)
confusionMatrixTrain <- table(loanstrain$Mod3ClassTrain, loanstrain$notfullypaid)
confusionMatrixTrain
```

```{r}
AccuracyTrain <- (7214+10)/(7214+16+1389+10)
AccuracyTrain
```

```{r}
loanstest$Mod3PredTest <- predict(redmod3, newdata = loanstest, type = "response")
loanstest$Mod3ClassTest <- ifelse(loanstest$Mod3PredTest > 0.5, 1,0)
confusionMatrixTest <- table(loanstest$Mod3ClassTest, loanstest$notfullypaid)
confusionMatrixTest
```

```{r}
AccuracyTest <- (814+1)/(814+1+142+11)
AccuracyTest
```

# Prediction on original data

```{r}
loans$Mod3Pred <- predict(redmod3, newdata = loans, type = "response")
loans$Mod3Class <- ifelse(loans$Mod3Pred > 0.5, 1,0)
confusionMatrix <- table(loans$Mod3Class, loans$notfullypaid)
confusionMatrix
```

```{r}
Accuracy <- (8028+17)/(8028+17+1522+11)
Accuracy
```

# New loans data with prob and class fields

```{r}
str(loans)
```

